### 2016-09-10 draft/draft.tex
Update latex draft

\usepackage[
backend=biber,
style=ieee,
sorting=ynt
]{biblatex}

@@

Having been a follower of the
bitcoin\href{https://bitcoin.org/en/}{(Bitcoin)} movement for more than
two years, the author has been looking for an appropriate opportunity to
apply the blockchain data structure, which is a key part of the
technology underlying bitcoin, to a project of his own. This project's
goal of verifiability of team work fits a blockchain's ability to
provide a distributed ledger of transactions or events.

The following points provide an overview of the author's personal aims
for the project: - Become familiar with native mobile app development by
utilising existing JavaScript knowledge in the context of React
Native\href{https://facebook.github.io/react-native/}{(RN)}. - Learn how
to write smart contracts in Solidity \emph{\href{}{REF}} that are useful
in the context of the project and for existing platforms such as
Ethereum \emph{\href{}{REF}}. - Implement a form of blockchain
technology, by learning how to set up, run \& maintain a a blockchain
proficiently. - Gain a fundamental understanding of Docker
\emph{\href{}{REF}} and how it leverages ``containerisation'' to achieve
straightforward deployments of development environments.

@@

The project followed the methodology of the Unified Process (henceforth
UP) framework for software
development\href{http://tocs.ulb.tu-darmstadt.de/82047766.pdf}{(Jacobson,
Ivar, et al. The unified software development process. Vol. 1. Reading:
Addison-wesley, 1999.)}, and was thus structured in terms of iterative
phases. The system was at first modeled as a whole by using a
theoretical deployment diagram in order to gain an understanding of
which technological entities were required to neatly encapsulate the
responsibilities within the system. Sequence diagrams were then used to
model more complex client-server-blockchain interactions during each UP
iteration, where a new stage usually took the form of defining and
implementing a new domain within the API, e.g.~task management or user
management.

Given the limited amount of time available for a project of these
dimensions, along with the research and experimentation required to
validate whether the described goals were even achievable, user
requirements were captured using a heuristic of what the essential
functionality would be for the client-side application, and thus the API
in general, in order for a user to create, manage and discuss their open
tasks with their team. This was done by utilising the MoSCoW
categorisation scheme \emph{\href{}{REF}}.

During the implementation stage of the project, a daily log (\textbf{see
Appendix X}) was kept of the key advances and issues encountered during
that day of development. This served as a useful tool to summarise the
progress made, as well as providing an immediate opportunity to document
issues that were encountered.\\
Throughout the implementation stage, required programming work was
broken down into separate units of work by using Github
Issues\href{}{(REF)} and organised into a Kanban board\href{}{(REF)}
with the help of waffle.io\href{https://waffle.io/}{(waffle.io)}.

@@

Additionally, a Gantt chart (\textbf{see Appendix X}) was established at
the outset of the project in order to visualise the progression of the
project over time and to ensure the amount of work set out to be
completed was within a realistic margin for the amount of time available
to the author.

@@

A 2004 study by the National Research Council showed that upwards of
40\% of high school students feel disengaged from learning and exert
little effort on school work
\href{http://cep-dc.org/displayDocument.cfm?DocumentID=405}{(Center on
Education Policy 2012)}. More worryingly still, 70\% of dropouts stated
lack of motivation, in a 2006 study, as the key reason for their
departure from school \href{}{(Bridgeland, J.M. et al., 2006. The Silent
Epidemic: Perspectives of High School Dropouts)}. In liaison with these
findings, a study which tracked the educational careers of a group of
individuals found that 23\% of those who dropped out cited a sense of
not belonging as the reason for their departure from school
\href{}{(Berktold et al., 1998, Table 6)}.

While the factors for a student's motivation (or lack of it) regarding
their education may stem from a myriad of factors, engaging the
individual on a social level may thus be posited as a key driver in
motivating them to remain engaged in the process \href{}{((Johnson et
al., 2001; Newmann, Wehlage, and Lamborn, 1992; Tinto, 1993; Wehlage,
Rutter, Smith, Lesko, and Fernandez, 1989))}. Of course, simply being
bound into a social context does not mean that the student will make
meaningful progress regarding their work, as attendance does not
constitute attention to what is being taught \href{}{(\emph{ibid})}.

@@

\begin{quote}
``Rewarding specific actions that students can control, such as
completing homework, yields better results than rewarding
accomplishments that may seem beyond their reach or out of their
control, such as whether they earn an A grade.''\\
\href{}{(Usher, Kober; Student Motivation - An overlooked piece of
school reform)}
\end{quote}

\textbf{Figure X} visualises a widely held theory regarding educational
conditions that promote intellectual engagement\href{}{(Committee on
Increasing High School Students' Engagement and Motivation to Learn \&
National Research Council 2003)}. The proposed system would thus provide
a shift in the educational context, providing a sense of control (teams
register their own tasks), values \& goals (wanting to help each other
and/or earn a higher ranking) and social connectedness (forming and
maintaining teams, reaching out for and offering help).

@@

\paragraph{\texorpdfstring{SocialX\protect\hyperlink{needsLink}{(Learning
from peers: motivating students through reputation systems )},
\protect\hyperlink{needsLink}{(Collaborative projects and self
evaluation within a social reputation-based exercise-sharing system
)}}{SocialX(Learning from peers: motivating students through reputation systems ), (Collaborative projects and self evaluation within a social reputation-based exercise-sharing system )}}\label{socialxlearning-from-peers-motivating-students-through-reputation-systems-collaborative-projects-and-self-evaluation-within-a-social-reputation-based-exercise-sharing-system}

SocialX is an exercise sharing tool which enables students to earn
reputation points, which are visible to their peers and the given
subject's teacher, by submitting solutions to exercises. A further
source of reputation are endorsements, which a student may receive from
their peers if they were inspired to reuse the student's solution.

@@

\paragraph{\texorpdfstring{WikiSpaces
Classroom\href{https://www.wikispaces.com/content/classroom/about}{(WikiSpaces)}}{WikiSpaces Classroom(WikiSpaces)}}\label{wikispaces-classroomwikispaces}

WikiSpaces is a web app which provides a virtual classroom workspace in
which teachers and students can work on written projects individually or
in teams. Features include a social media-like feed, collaborative
writing and commenting tools, and a project management system.

@@

\paragraph{\texorpdfstring{HaikuLearning\href{https://www.haikulearning.com/}{(HaikuLearning)}}{HaikuLearning(HaikuLearning)}}\label{haikulearninghaikulearning}

HaikuLearning, similarly to WikiSpaces, is a web app which attempts to
emulate and enhance the classroom experience for both school teachers
and students. It offer a unified environment to hand in assignments and
to receive feedback \& grades, as well as enabling integrations with
Google Apps for extensibility.

@@

Before proceeding with the following section, two frequently used
concepts should be clearly delineated: blockchains and smart contracts.
A blockchain is a data structure composed of a list of blocks of
transactions, where each block is linked back to the previous block in
the chain\href{}{(Mastering Bitcoin)}. Each block is identified by a
hash, which is stored in the block's header alongside the previous
block's hash, known as the parent block(\emph{ibid}), thus providing a
traceable, reverse-chronological chain of transactions over time.\\
The concept of a smart contract was originally developed by Nick Szabo
in 1997, and is usually evoked in terms of a digital protocol which
enforces and/or verifies the performance of a contract between two or
more
parties\href{http://szabo.best.vwh.net/smart_contracts_idea.html}{(smart\_contracts\_idea)}.

At the outset of the project, a key decision which had to be made was
the choice of blockchain implementation to be used. This decision was
critical, as discovering a major flaw in the implementation or an
incompatibility between what the implementation could provide and what
was needed, would have resulted in (at least temporary) deadlock for the
entire project.\\
The author's initial experimentation with
OpenChain\href{https://www.openchain.org/}{(OpenChain)}, which seemed to
offer the possibly useful ability to chain contracts to each other, was
quickly abandoned as the available documentation was insufficiently
detailed and partially ambiguous. This would have left the project open
to a lot of guesswork and thus the above mentioned risk of deadlock.\\
MultiChain\href{http://www.multichain.com/}{(MultiChain)}, on the other
hand, was an implementation which had well-defined documentation, but it
assumed a level of pre-existing familiarity with the API of a
blockchain, as well as providing preciously little context as to how
smart contracts could be developed for the platform in an effective
way.\\
The apparent drawbacks of the two mentioned blockchain implementations
therefore led the author to the
Eris\href{https://erisindustries.com/}{(Eris)} platform. Eris itself
provides a suite of tools which wrap and augment a
Tendermint\href{http://tendermint.com/}{(Tendermint)} blockchain
implementation. Eris's wealth of documentation and its well-structured
command line interface (henceforth CLI) were the deciding factors for it
becoming the blockchain implementation of choice. The documentation
provided not only step-by-step examples of how a developer could
configure and deploy a blockchain, but also provided in-depth tutorials
and examples on how to write \& deploy
Solidity\href{http://solidity.readthedocs.io/en/latest/}{(Solidity)}
smart contracts for the platform; highly valuable for a smart contract
novice such as the author. Furthermore, the fact that Tendermint is an
entirely separate software project meant that there was an additional
repository of documentation for this specific type of blockchain and the
computer science theory backing it.

@@

Considering the author's background as a JavaScript developer, the
natural choice for a web server for the project was
NodeJS\href{https://nodejs.org/en/}{(NodeJS)}. NodeJS provides a great
level of flexibility and extensibility thanks to its huge ecosystem of
open-source libraries available through its package manager,
NPM\href{https://www.npmjs.com/}{(npm)}, thus providing an essential
basis to help mitigate the possibly complicated technical details of
interfacing with a low-level blockchain.\\
As JavaScript is an ever-evolving language, a choice had to be made as
to whether the commonly supported
ECMAScript\href{http://www.ecma-international.org/}{(ECMAScript)} 5
(ES5) specification should be used or the modern standard, ES6. ES5 was
selected to provide the maximum level of compatibility for wherever the
server is being deployed, as native support for ES6 is not widespread as
of the point of writing and transpiling the server's code from ES6 to
ES5 for each deployment would have been a lot more trouble than use for
a humble web server.

@@

Express\href{https://expressjs.com/}{(Express)} is a minimalist
framework for NodeJS web servers which provides a powerful layer of
abstraction on top of NodeJS's raw server interface. Express helped
accelerate the development of the API, while leaving the possibility of
fine-tuning the NodeJS server itself intact.

@@

The eris-wrapper \& eris-logger modules were adopted from a ``Hello
World''-style example Eris provides to showcase how a NodeJS server
implementing their platform could be
structured\href{https://github.com/eris-ltd/hello-eris}{(Hello Eris)}.\\
eris-wrapper provides a convenient abstraction of low-level bindings
that have to take place between the NodeJS server and the Tendermint
blockchain, while eris-logger simply sets up a simple logger with
different logging types, such as \texttt{ERROR}, \texttt{INFO} and
\texttt{DEBUG}.

@@

Async\href{https://caolan.github.io/async/}{(Async)} is a utility
library that helps manage the flow of asynchronous functions, which are
a common occurrence in NodeJS. Within the scope of the server, Async was
primarily used to chain together multiple sequential interactions with
the blockchain.

@@

During the MSc's GC02 ``App Design'' course, the author and his team had
the opportunity to build a React web application which was also required
to run as a mobile application. This was achieved by wrapping the web
app with the Apache Cordova\href{https://cordova.apache.org/}{(Cordova)}
mobile development framework. This experience revealed two key insights
regarding the intersection between web- and mobile-based client-side
development:\\
Firstly, that Facebook's
React\href{https://facebook.github.io/react/}{(React)} web development
library, in liaison with Dan Abramov's
Redux\href{http://redux.js.org/}{(Redux)} library for state management,
could be applied almost seamlessly to the context of a mobile
application. React's philosophy of thinking in terms of individual view
components, together with Redux's implementation-agnostic approach to
managing the application's state, largely abstracted away the conceptual
differences between building an HTML document and building a mobile app
view.\\
Secondly, that porting a web app to a mobile app simply does not provide
an authentic native experience on a mobile device, as it is almost
impossible to account for and implement all the differences and
individual nuances in, for example, animation styles between mobile
operating systems. This resulted in the look \& feel of the application
being closer to that of a mobile browser rather than a native mobile
app.

Based on these two lessons learnt, the author decided to use Facebook's
React Native\href{https://facebook.github.io/react-native/}{(React
Native)} for the project, which grafts a truly native mobile development
framework on top of the outstanding React library, along with the now
familiar Redux library.\\
React Native offers all the benefits of React (high levels of modularity
\& encapsulation, composability) while allowing the developer to hook
into events and effects of the native mobile operating system. This
means that transitioning between views smoothly or sending a push
notification, for example, becomes a trivial undertaking.\\
Redux provides a well-defined interface to manage the state of the
application's data by managing all of it through a centralised
\texttt{store} object, the contents of which can be altered exclusively
by a set of actions previously defined by the developer. This provides a
level of clarity of how the application's state mutates over time that
is hard to achieve in a traditional Model-View-Controller\href{}{(REF)}
(henceforth MVC) approach, in which many controllers have access to the
application's state simultaneously.

@@

redux-thunk\href{https://github.com/gaearon/redux-thunk}{(redux-thunk)}
is a helper library which enables the use of asynchronous functions as
the aforementioned actions that affect the application's state. This is
essential to allow control over how and when the client application
sends data to the blockchain, and how it fetches data from the
blockchain to then integrate it into its local state.

@@

redux-logger\href{https://github.com/evgenyrodionov/redux-logger}{(redux-logger)}
supplements Redux itself by automatically logging actions and the state
changes they trigger to the development console. This meant a
significant amount of time could be saved during the project by avoiding
the need for a custom logger implementation, or worse yet, littering the
codebase with sporadic logging statements.

@@

tcomb-form-native\href{https://github.com/gcanti/tcomb-form-native}{(tcomb-form-native)}
provides a framework for forms and form validation in React Native. As
form validation in particular can take up an inordinate amount of
development time, this library was essential in order to keep the
project on track given the severe time constraints in relation to its
size.

@@

Mocha\href{https://mochajs.org/}{(Mocha)} was chosen as the testing
framework for the project due to its specific aim of simplifying the
testing of asynchronous JavaScript code. This clearly applied to the
needs of the planned API which would be largely built on asynchronous
functions.

@@

Chai\href{http://chaijs.com/}{(Chai)} is an assertion library for both
test-driven development (TDD) and behaviour-driven development (BDD).
More specifically, Chai's TDD \texttt{assert} function was used to keep
test assertions simple and testable against an expected result.

@@

Istanbul\href{https://gotwarlost.github.io/istanbul/}{(Istanbul)} is a
code coverage tool which was used to gain an overview of how much of the
API was currently covered by tests. This occurred via coverage reports,
which were generated after every run of the test suite.

@@

GanttPro\href{https://ganttpro.com/}{(GanttPro)}, an online creator and
editor for Gantt charts, was used to create a timeline for the project,
while Google Docs\href{https://www.google.com/docs/about/}{(GoogleDocs)}
in conjunction with draw.io\href{https://www.draw.io/}{(draw.io)} were
used to capture requirements and create UML diagrams.\\
This report was originally written in markdown and transposed to the
Latex format using pandoc\href{http://pandoc.org/}{(pandoc)}.

@@

In order to avoid switching between editors to get the best development
support for disparate parts of the project's code, Github's
Atom\href{https://atom.io/}{(Atom)} editor was chosen due to its level
of customisability and huge selection of plugins. Atom was particularly
suited for writing React Native code thanks to Facebook's
Nuclide\href{https://nuclide.io/}{(Nuclide)} plugin, which enables Atom
to approximate a richness of features typically only found in an
Integrated Development Environment (IDE), by offering an in-built
debugger, code snippets, and Facebook's own static type analyser which
is discussed below.

@@

Flowtype\href{https://www.flowtype.org/}{(Flowtype)}, a static type
analyser for JavaScript, is a further in-built feature of the Nuclide
plugin. Flowtype allows the developer to define a type for a variable, a
type signature for a function, and even to create custom union \&
intersection types. Flowtype then checks whether the defined type
specifications are adhered to and warns if it detects a TypeError. This
is an incredibly useful tool for a dynamically-typed language such as
JavaScript, where unintended type coercion is a common ailment.

@@

ESLint\href{http://eslint.org/}{(ESLint)} was chosen as the linter for
both the React Native app and the NodeJS server, due to its support for
detecting syntactical \& stylistic errors in both regular JavaScript and
JSX, a mix of JavaScript and HTML used by Facebook for React.\\
For the smart contracts, written in Solidity, the
\texttt{solidity-linter}\href{https://atom.io/packages/linter-solidity}{(solidity-linter)}
Atom plugin was used, which avoided a significant amount of frustration
in comparison to waiting for errors to be detected at compile-time by
the Solidity compiler built into Eris's tooling.

@@

Debugging within the project was performed in three different ways due
to the variability of environments within the technology stack. For the
React Native app, Google Chrome's
DevTools\href{https://developers.google.com/web/tools/chrome-devtools/?hl=en}{(devTools)}
were utilised alongside the Xcode iOS
simulator\href{https://developer.apple.com/library/ios/documentation/IDEs/Conceptual/iOS_Simulator_Guide/Introduction/Introduction.html}{(iOS
Simulator)} in order to interact with the app and receive log outputs
side-by-side in real-time.\\
For the NodeJS server, the aforementioned \texttt{eris-logger} module
was used to log API-related logging statements to the terminal.\\
At the bare-metal level, Eris provided a way of continuously logging the
activity of the Tendermint blockchain. Due to all operations
approximating those of assembly language and thus being represented in
hexadecimal, this was not useful in terms of locating bugs, but it
provided a sanity check to ensure that the chain was performing the
expected operations when instructed to do so by the NodeJS server,
excluding a possible cause if a bug was being searched for.

@@

The \texttt{git}\href{https://git-scm.com/}{(git)} command line utility
was used to manage the codebase's development over time, in conjunction
with GitHub\href{https://github.com/}{(GitHub)} to provide a remote
backup. For more involved \texttt{git} operations, such as merging
branches, Atlassian's
SourceTree\href{https://www.atlassian.com/software/sourcetree}{(SourceTree)}
GUI was used to avoid mistyping complex terminal commands and thus
potentially performing unwanted changes to the version history.

@@

Docker\href{https://www.docker.com/}{(Docker)} is a platform which
allows the creation of self-contained virtual environments for software
development, to mitigate arbitrary local differences between development
environments. It was an essential part of the development process as
Eris's tooling leverages Docker heavily to deploy blockchain
instances.\\
Bash shell scripts (\textbf{see Appendix X}) were constructed by the
author in order to automate processes such as hydrating the local
terminal environment with variables required to run the blockchain, or
booting the Docker virtual machine and the local blockchain instance.

@@

In a typical software engineering project, there is a need for strong
levels of communication between all the stakeholders involved in the
endeavour, in order to ensure all the needs the software should meet are
in fact met \emph{\href{}{(REF)}}. A frequent phenomenon within these
interactions is that requirements are revealed to the development team
in a haphazard way or are simply misinterpreted by the team
\emph{\href{}{(REF?)}}.\\
Fortunately these common pitfalls were a moot point within this project,
as the act of establishing requirements was entirely focused around the
question: ``What functionality does the system require, at a minimum, to
fulfill its stated aims?''. This provided a clear focus on what was
absolutely needed for a minimum viable product (henceforth
MVP)\href{https://www.techopedia.com/definition/27809/minimum-viable-product-mvp}{(MVP;
better ref)} that an individual and/or group of individuals could use in
a meaningful way. This meant both functional and non-functional
requirements were focused around three domains: tasks, users, and teams.
The requirements were prioritised according to the MoSCoW
system\href{https://www.dsdm.org/content/moscow-prioritisation}{(MoSCoW)},
with requirements being ranked from ``Must Have'' through ``Should
Have'' and ``Could Have'', with the final category being ``Won't Have
(in this development cycle)''. The project's ``Must Have'' requirements
had to be strictly limited to what was viable within the project's
timeframe, as there were a large number of known unknowns (e.g.~the
Solidity language) and unknown unknowns (unforeseeable issues with the
API, blockchain or development environment). The ``Should Have'' and
``Could Have'' categories therefore largely express targets for more
sophisticated future iterations of the system.

@@

Use cases were constructed in parallel with the initial UI sketches,
thus helping to visualise the flow of user-system interactions.\\
\textbf{Figure xx} below represents an overview of the uses cases that
were constructed, while the detailed use cases can be reviewed in
\textbf{Appendix X}.

@@

>>>>

\usepackage[
backend=biber,
style=ieee
]{biblatex}

@@

Having been a follower of the
bitcoin\cite{1bitcoin} movement for more than
two years, the author has been looking for an appropriate opportunity to
apply the blockchain data structure, which is a key part of the
technology underlying bitcoin, to a project of his own. This project's
goal of verifiability of team work fits a blockchain's ability to
provide a distributed ledger of transactions or events.

The following points provide an overview of the author's personal aims
for the project: - Become familiar with native mobile app development by
utilising existing JavaScript knowledge in the context of React
Native\cite{2react-native}. - Learn how
to write smart contracts in Solidity\cite{1solidity} that are useful
in the context of the project and for existing platforms such as
Ethereum\cite{1ethereum}. - Implement a form of blockchain
technology, by learning how to set up, run \& maintain a a blockchain
proficiently. - Gain a fundamental understanding of Docker\cite{1docker} and how it leverages ``containerisation'' to achieve
straightforward deployments of development environments.

@@

The project followed the methodology of the Unified Process (henceforth
UP) framework for software
development\cite{jacobson1999unified}, and was thus structured in terms of iterative
phases. The system was at first modeled as a whole by using a
theoretical deployment diagram in order to gain an understanding of
which technological entities were required to neatly encapsulate the
responsibilities within the system. Sequence diagrams were then used to
model more complex client-server-blockchain interactions during each UP
iteration, where a new stage usually took the form of defining and
implementing a new domain within the API, e.g.~task management or user
management.

Given the limited amount of time available for a project of these
dimensions, along with the research and experimentation required to
validate whether the described goals were even achievable, user
requirements were captured using a heuristic of what the essential
functionality would be for the client-side application, and thus the API
in general, in order for a user to create, manage and discuss their open
tasks with their team. This was done by utilising the MoSCoW
categorisation scheme\cite{1moscow}.

During the implementation stage of the project, a daily log was kept of the key advances and issues encountered during
that day of development. This served as a useful tool to summarise the
progress made, as well as providing an immediate opportunity to document
issues that were encountered.\\
Throughout the implementation stage, required programming work was
broken down into separate units of work by using Github
Issues\cite{1githubissues} and organised into a Kanban board\href{}{(REF)}
with the help of waffle.io\cite{1waffle}.

@@

@@

A 2004 study by the National Research Council showed that upwards of
40\% of high school students feel disengaged from learning and exert
little effort on school work\cite{Usher2013}. More worryingly still, 70\% of dropouts stated
lack of motivation, in a 2006 study, as the key reason for their
departure from school\cite{bridgeland2006silent}. In liaison with these
findings, a study which tracked the educational careers of a group of
individuals found that 23\% of those who dropped out cited a sense of
not belonging as the reason for their departure from school\cite{berktold1998subsequent}.

While the factors for a student's motivation (or lack of it) regarding
their education may stem from a myriad of factors, engaging the
individual on a social level may thus be posited as a key driver in
motivating them to remain engaged in the process\cite{johnson2001students}. Of course, simply being
bound into a social context does not mean that the student will make
meaningful progress regarding their work, as attendance does not
constitute attention to what is being taught.

@@

\begin{quote}
``Rewarding specific actions that students can control, such as
completing homework, yields better results than rewarding
accomplishments that may seem beyond their reach or out of their
control, such as whether they earn an A grade.''\cite{Usher2013}
\end{quote}

Figure 2.1 visualises a widely held theory regarding educational
conditions that promote intellectual engagement\cite{CommitteeonIncreasingHighSchoolStudentsEngagementandMotivationtoLearn2003}. The proposed system would thus provide
a shift in the educational context, providing a sense of control (teams
register their own tasks), values \& goals (wanting to help each other
and/or earn a higher ranking) and social connectedness (forming and
maintaining teams, reaching out for and offering help).

@@

\paragraph{SocialX}\label{socialx}

SocialX\cite{Temperini2008}\cite{Sterbini2009} is an exercise sharing tool which enables students to earn
reputation points, which are visible to their peers and the given
subject's teacher, by submitting solutions to exercises. A further
source of reputation are endorsements, which a student may receive from
their peers if they were inspired to reuse the student's solution.

@@

\paragraph{WikiSpaces}\label{wikispaces}

WikiSpaces\cite{1wikispaces} is a web app which provides a virtual classroom workspace in
which teachers and students can work on written projects individually or
in teams. Features include a social media-like feed, collaborative
writing and commenting tools, and a project management system.

@@

\paragraph{HaikuLearning}\label{haikulearning}

HaikuLearning\cite{haikulearning}, similarly to WikiSpaces, is a web app which attempts to
emulate and enhance the classroom experience for both school teachers
and students. It offer a unified environment to hand in assignments and
to receive feedback \& grades, as well as enabling integrations with
Google Apps for extensibility.

@@

Before proceeding with the following section, two frequently used
concepts should be clearly delineated: blockchains and smart contracts.
A blockchain is a data structure composed of a list of blocks of
transactions, where each block is linked back to the previous block in
the chain\cite{antonopoulos2014mastering}. Each block is identified by a
hash, which is stored in the block's header alongside the previous
block's hash, known as the parent block\cite{antonopoulos2014mastering}, thus providing a
traceable, reverse-chronological chain of transactions over time.\\
The concept of a smart contract was originally developed by Nick Szabo
in 1997, and is usually evoked in terms of a digital protocol which
enforces and/or verifies the performance of a contract between two or
more
parties\cite{1szabo}.

At the outset of the project, a key decision which had to be made was
the choice of blockchain implementation to be used. This decision was
critical, as discovering a major flaw in the implementation or an
incompatibility between what the implementation could provide and what
was needed, would have resulted in (at least temporary) deadlock for the
entire project.\\
The author's initial experimentation with
OpenChain\cite{1openchain}, which seemed to
offer the possibly useful ability to chain contracts to each other, was
quickly abandoned as the available documentation was insufficiently
detailed and partially ambiguous. This would have left the project open
to a lot of guesswork and thus the above mentioned risk of deadlock.\\
MultiChain\cite{1multichain}, on the other
hand, was an implementation which had well-defined documentation, but it
assumed a level of pre-existing familiarity with the API of a
blockchain, as well as providing preciously little context as to how
smart contracts could be developed for the platform in an effective
way.\\
The apparent drawbacks of the two mentioned blockchain implementations
therefore led the author to the
Eris\cite{1erisindustries} platform. Eris itself
provides a suite of tools which wrap and augment a
Tendermint\cite{1tendermint} blockchain
implementation. Eris's wealth of documentation and its well-structured
command line interface (henceforth CLI) were the deciding factors for it
becoming the blockchain implementation of choice. The documentation
provided not only step-by-step examples of how a developer could
configure and deploy a blockchain, but also provided in-depth tutorials
and examples on how to write \& deploy
Solidity\cite{1solidity}
smart contracts for the platform; highly valuable for a smart contract
novice such as the author. Furthermore, the fact that Tendermint is an
entirely separate software project meant that there was an additional
repository of documentation for this specific type of blockchain and the
computer science theory backing it.

@@

Considering the author's background as a JavaScript developer, the
natural choice for a web server for the project was
NodeJS\cite{1node}. NodeJS provides a great
level of flexibility and extensibility thanks to its huge ecosystem of
open-source libraries available through its package manager,
NPM\cite{1npm}, thus providing an essential
basis to help mitigate the possibly complicated technical details of
interfacing with a low-level blockchain.\\
As JavaScript is an ever-evolving language, a choice had to be made as
to whether the commonly supported
ECMAScript\cite{1ecma} 5
(ES5) specification should be used or the modern standard, ES6. ES5 was
selected to provide the maximum level of compatibility for wherever the
server is being deployed, as native support for ES6 is not widespread as
of the point of writing and transpiling the server's code from ES6 to
ES5 for each deployment would have been a lot more trouble than use for
a humble web server.

@@

Express\cite{1express} is a minimalist
framework for NodeJS web servers which provides a powerful layer of
abstraction on top of NodeJS's raw server interface. Express helped
accelerate the development of the API, while leaving the possibility of
fine-tuning the NodeJS server itself intact.

@@

The eris-wrapper \& eris-logger modules were adopted from a ``Hello
World''-style example Eris provides to showcase how a NodeJS server
implementing their platform could be
structured\cite{1helloeris}.\\
eris-wrapper provides a convenient abstraction of low-level bindings
that have to take place between the NodeJS server and the Tendermint
blockchain, while eris-logger simply sets up a simple logger with
different logging types, such as \texttt{ERROR}, \texttt{INFO} and
\texttt{DEBUG}.

@@

Async\cite{1async} is a utility
library that helps manage the flow of asynchronous functions, which are
a common occurrence in NodeJS. Within the scope of the server, Async was
primarily used to chain together multiple sequential interactions with
the blockchain.

@@

During the MSc's GC02 ``App Design'' course, the author and his team had
the opportunity to build a React web application which was also required
to run as a mobile application. This was achieved by wrapping the web
app with the Apache Cordova\cite{1cordova}
mobile development framework. This experience revealed two key insights
regarding the intersection between web- and mobile-based client-side
development:\\
Firstly, that Facebook's
React\cite{1react} web development
library, in liaison with Dan Abramov's
Redux\cite{1redux} library for state management,
could be applied almost seamlessly to the context of a mobile
application. React's philosophy of thinking in terms of individual view
components, together with Redux's implementation-agnostic approach to
managing the application's state, largely abstracted away the conceptual
differences between building an HTML document and building a mobile app
view.\\
Secondly, that porting a web app to a mobile app simply does not provide
an authentic native experience on a mobile device, as it is almost
impossible to account for and implement all the differences and
individual nuances in, for example, animation styles between mobile
operating systems. This resulted in the look \& feel of the application
being closer to that of a mobile browser rather than a native mobile
app.

Based on these two lessons learnt, the author decided to use Facebook's
React Native\cite{2react-native} for the project, which grafts a truly native mobile development
framework on top of the outstanding React library, along with the now
familiar Redux library.\\
React Native offers all the benefits of React (high levels of modularity
\& encapsulation, composability) while allowing the developer to hook
into events and effects of the native mobile operating system. This
means that transitioning between views smoothly or sending a push
notification, for example, becomes a trivial undertaking.\\
Redux provides a well-defined interface to manage the state of the
application's data by managing all of it through a centralised
\texttt{store} object, the contents of which can be altered exclusively
by a set of actions previously defined by the developer. This provides a
level of clarity of how the application's state mutates over time that
is hard to achieve in a traditional Model-View-Controller\cite{krasner1988description}
(henceforth MVC) approach, in which many controllers have access to the
application's state simultaneously.

@@

redux-thunk\cite{1thunk}
is a helper library which enables the use of asynchronous functions as
the aforementioned actions that affect the application's state. This is
essential to allow control over how and when the client application
sends data to the blockchain, and how it fetches data from the
blockchain to then integrate it into its local state.

@@

redux-logger\cite{1reduxlogger}
supplements Redux itself by automatically logging actions and the state
changes they trigger to the development console. This meant a
significant amount of time could be saved during the project by avoiding
the need for a custom logger implementation, or worse yet, littering the
codebase with sporadic logging statements.

@@

tcomb-form-native\cite{1tcombform}
provides a framework for forms and form validation in React Native. As
form validation in particular can take up an inordinate amount of
development time, this library was essential in order to keep the
project on track given the severe time constraints in relation to its
size.

@@

Mocha\cite{1mocha} was chosen as the testing
framework for the project due to its specific aim of simplifying the
testing of asynchronous JavaScript code. This clearly applied to the
needs of the planned API which would be largely built on asynchronous
functions.

@@

Chai\cite{1chai} is an assertion library for both
test-driven development (TDD) and behaviour-driven development (BDD).
More specifically, Chai's TDD \texttt{assert} function was used to keep
test assertions simple and testable against an expected result.

@@

Istanbul\cite{1istanbul} is a
code coverage tool which was used to gain an overview of how much of the
API was currently covered by tests. This occurred via coverage reports,
which were generated after every run of the test suite.

@@

GanttPro\cite{1ganttpro}, an online creator and
editor for Gantt charts, was used to create a timeline for the project,
while Google Docs\cite{2googledocs}
in conjunction with draw.io\cite{1drawio} were
used to capture requirements and create UML diagrams.\\
This report was originally written in markdown and transposed to the
Latex format using pandoc\cite{1pandoc}.

@@

In order to avoid switching between editors to get the best development
support for disparate parts of the project's code, Github's
Atom\cite{1atom} editor was chosen due to its level
of customisability and huge selection of plugins. Atom was particularly
suited for writing React Native code thanks to Facebook's
Nuclide\cite{1nuclide} plugin, which enables Atom
to approximate a richness of features typically only found in an
Integrated Development Environment (IDE), by offering an in-built
debugger, code snippets, and Facebook's own static type analyser which
is discussed below.

@@

Flowtype\cite{1flowtype}, a static type
analyser for JavaScript, is a further in-built feature of the Nuclide
plugin. Flowtype allows the developer to define a type for a variable, a
type signature for a function, and even to create custom union \&
intersection types. Flowtype then checks whether the defined type
specifications are adhered to and warns if it detects a TypeError. This
is an incredibly useful tool for a dynamically-typed language such as
JavaScript, where unintended type coercion is a common ailment.

@@

ESLint\cite{1eslint} was chosen as the linter for
both the React Native app and the NodeJS server, due to its support for
detecting syntactical \& stylistic errors in both regular JavaScript and
JSX, a mix of JavaScript and HTML used by Facebook for React.\\

@@

Debugging within the project was performed in three different ways due
to the variability of environments within the technology stack. For the
React Native app, Google Chrome's
DevTools\cite{1chromedevtools}
were utilised alongside the Xcode iOS
simulator\cite{1iossimulator} in order to interact with the app and receive log outputs
side-by-side in real-time.\\
For the NodeJS server, the aforementioned \texttt{eris-logger} module
was used to log API-related logging statements to the terminal.\\
At the bare-metal level, Eris provided a way of continuously logging the
activity of the Tendermint blockchain. Due to all operations
approximating those of assembly language and thus being represented in
hexadecimal, this was not useful in terms of locating bugs, but it
provided a sanity check to ensure that the chain was performing the
expected operations when instructed to do so by the NodeJS server,
excluding a possible cause if a bug was being searched for.

@@

The \texttt{git}\cite{1git} command line utility
was used to manage the codebase's development over time, in conjunction
with GitHub\cite{1github} to provide a remote
backup. For more involved \texttt{git} operations, such as merging
branches, Atlassian's
SourceTree\cite{1sourcetree}
GUI was used to avoid mistyping complex terminal commands and thus
potentially performing unwanted changes to the version history.

@@

Docker\cite{1docker} is a platform which
allows the creation of self-contained virtual environments for software
development, to mitigate arbitrary local differences between development
environments. It was an essential part of the development process as
Eris's tooling leverages Docker heavily to deploy blockchain
instances.\\
Bash shell scripts (\textbf{see Appendix X}) were constructed by the
author in order to automate processes such as hydrating the local
terminal environment with variables required to run the blockchain, or
booting the Docker virtual machine and the local blockchain instance.

\clearpage

@@

The act of establishing requirements was entirely focused around the
question: ``What functionality does the system require, at a minimum, to
fulfill its stated aims?''. This provided a clear focus on what was
absolutely needed for a minimum viable product (henceforth
MVP)\cite{ries2009minimum} that an individual and/or group of individuals could use in
a meaningful way. This meant both functional and non-functional
requirements were focused around three domains: tasks, users, and teams.
The requirements were prioritised according to the MoSCoW
system\cite{1moscow},
with requirements being ranked from ``Must Have'' through ``Should
Have'' and ``Could Have'', with the final category being ``Won't Have
(in this development cycle)''. The project's ``Must Have'' requirements
had to be strictly limited to what was viable within the project's
timeframe, as there were a large number of known unknowns (e.g.~the
Solidity language) and unknown unknowns (unforeseeable issues with the
API, blockchain or development environment). The ``Should Have'' and
``Could Have'' categories therefore largely express targets for more
sophisticated future iterations of the system.

@@

Use cases were constructed in parallel with the initial UI sketches,
thus helping to visualise the flow of user-system interactions.\\
Figure 3.4 below represents an overview of the uses cases that
were constructed, while the detailed use cases can be reviewed in
\textbf{Appendix X}.

@@

\printbibliography
